Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices

### 2022-07-07 paper.tex
[add] DoS updates

%-------------------------------------------------------------------------- sec
\section{\uppercase{Denial of Service}}
\label{sec:dos}
%\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
attacks and delegates the responsibility to handle this to the application.
%
\mEdhoc{} itself has limited functionality, few negotiable parameters and
processing rules.
%
Hence, we argue in this section that its attack surface
is limited from a DoS perspective.
%
However, \mEdhoc{} leaves much of the semantics of its messages undefined.
%
This increases the risk of applications vulnerable to DoS via \mEdhoc{}, even
though \mEdhoc{} itself is arguably not particularly weak in this respect.
%

@@

No recommendation is given on how to minimize DoS effects stemming from use of
error messages, which can be used in an arbitrary way in \mEdhoc{}.
%
We now elaborate on some considerations application developers needs to take
into account.
%

\paragraph{Amplification and asymmetry.}
For the same reasons as given in the IKEv2 DDoS analysis
by~\cite{rfc8019} the responder is more vulnerable than the
initiator.
%
The key similarity between IKEv2 and \mEdhoc{} here is that the initiator is
anonymous in the first message and that creating the first message is
computationally cheap and requires no storage.
%
No element in the first \mEdhoc{} message can be verified by the receiver as
being legitimate as long as the format is correct and elliptic curve points
lie on the proposed curve.
%
\mEdhoc{} leave it to the application and its
transport layer to resolve any issues.
%
In addition to the computational resources the responder spends, communication
patterns in the underlying protocols may require the responder to
obtain and verify certificate chains, and verify
certificate-revocation servers as a result of the spoofed initial message.
%
For example, the content of the \mADone{} information element may hint the
responder to take further actions.
%

@@

Implementers are well advised to consult \cite{rfc8019}, which provides much
good general information about DoS for protocols with this structure.
%
However, ultimately, a product-specific risk assessment is required to
determine appropriate DDoS mitigations.
%

\paragraph{Reflection Attacks}
Reflection attacks are attacks where an adversary spoofs the source address of a
victim in a message and thereby tricks the responder to send a message to the
victim.
%
\mEdhoc{} leaves protection against this to the transport protocol and
application, and suggests using a return path reachability method similar to IKEv2.
%
We note that such reachability mechanisms can easily be used to make a responder
send reachability requests to arbitrary addressable targets.
%
Depending on the design of such reachability mechanisms, they may require that
unsolicited reachability requests are discarded, limiting the effects of
anonymous DoS attacks like these.
%

@@

\mPoint{In a constrained device randomness may be easily depleted}
Assume an adversary generate (deterministic) bogus message 1s and send these at
a sufficiently high rate to the responder.
%
\mEdhoc{} specifically targets constrained devices, that presumably may not
always have good randomness sources.
%
Is it possible to deplete or significantly reduce the entropy of the source and
in this way be able to predict the responder's random input to other sessions
(with other initiators)?
%
The LAKE WG have been talking about effects of poor randomness on signatures, so
this should be an issue, no?
%

\mPoint{\mEdhoc{} can deadlock if not used correctly}
An adversary may send forged error messages, and specifically reject proposed
ciphersuites.
%
The initiator is recommended to not try that ciphersuite with the responder
again in a new session for a while.
%
If the adversary rejects all options for a initiator this way, it can prevent
communication between the initiator and that responder until the cache is
cleared.
%
Potentially, an adversary could lock a party of a group out of communication all
together this way.
%
This needs to be considered by applications.
%

@@

%-------------------------------------------------------------------------- sec
\section{\uppercase{Discussion}}
\label{sec:discussion}
There are a few places where \mEdhoc{} can be improved,
which we found during this work and communicated to the authors.
%
We discuss them below.
%

>>>>

%-------------------------------------------------------------------------- sec
\section{\uppercase{Denial of Service}}
\label{sec:dos}
%\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
attacks and delegates the responsibility to handle this to the application.
%
The core \mEdhoc{} specification has limited functionality, few negotiable
parameters and processing rules.
%
Hence, we argue in this section that its attack surface
is limited from a DoS perspective.
%
However, \mEdhoc{} leaves much of the semantics of its messages undefined or
specified as (optional) extensions.
%
It is therefore difficult to draw the line between what is \mEdhoc{} and what is
an extension.
%
For example, the core specification does not include a
message transport mechanism.
%
This increases the risk of applications vulnerable to DoS via \mEdhoc{}, even
though \mEdhoc{} itself is arguably not particularly weak in this respect.
%

@@

No recommendation is given on how to minimize DoS effects stemming from use of
error messages, which can be used in an arbitrary way in \mEdhoc{}.
%
While there is some work on DDoS aspects of \mCoap{}, and much on IoT in
general, this does not capture the particular consequences originating with
\mEdhoc{}.
%
We now elaborate on some considerations application developers needs to take
into account.
%

\paragraph{Amplification and role asymmetry.}
For the same reasons as given in the IKEv2 DDoS analysis
by~\cite{rfc8019} the responder is more vulnerable than the
initiator.
%
The key similarity between IKEv2 and \mEdhoc{} here is that the initiator is
anonymous in the first message and that creating the first message is
computationally cheap and requires no storage.
%
No element in the first \mEdhoc{} message can be verified by the receiver as
being legitimate as long as the format is correct and elliptic curve points
lie on the proposed curve.
%
\mEdhoc{} leave it to the application and its
transport layer to resolve any issues.
%
In addition to the computational resources the responder spends, \mEdhoc{}
execution may require the responder to
obtain and verify certificate chains, and verify
certificate-revocation servers as a result of the spoofed initial message.
%
For example, the content of the \mADone{} information element may hint the
responder to take further actions.
%

@@

\paragraph{Reflection Attacks.}
Reflection attacks are attacks where an adversary spoofs the source address of a
victim in a message and thereby tricks the responder to send a message to the
victim.
%
\mEdhoc{} leaves protection against this to the transport protocol and
application, and suggests using a return path reachability method similar to IKEv2.
%
We note that such reachability mechanisms can easily be used to make a responder
send reachability requests to arbitrary addressable targets.
%
Depending on the design of such reachability mechanisms, they may require that
unsolicited reachability requests are discarded, limiting the effects of
anonymous DoS attacks like these.
%

@@

%\mPoint{In a constrained device randomness may be easily depleted}
%Assume an adversary generate (deterministic) bogus message 1s and send these at
%a sufficiently high rate to the responder.
%%
%\mEdhoc{} specifically targets constrained devices, that presumably may not
%always have good randomness sources.
%%
%Is it possible to deplete or significantly reduce the entropy of the source and
%in this way be able to predict the responder's random input to other sessions
%(with other initiators)?
%%
%The LAKE WG have been talking about effects of poor randomness on signatures, so
%this should be an issue, no?
%%

\mPoint{\mEdhoc{} can deadlock if not used correctly}

An adversary may send forged error messages, and specifically reject proposed
ciphersuites.
%
The initiator is recommended to not try that ciphersuite with the responder
again in a new session for a while.
%
If the adversary rejects all options for a initiator this way, it can prevent
communication between the initiator and that responder until the cache is
cleared.
%
Potentially, an adversary could lock a party of a group out of communication all
together this way.
%
This needs to be considered by applications.
%

@@

Implementers are well advised to consult \cite{rfc8019}, which provides much
good general information about DoS for key establishment
protocols with the 3-message structure employed by \mEdhoc{}.
%
However, ultimately, a product-specific risk assessment is required to
determine appropriate DDoS mitigations.
%

\section{\uppercase{Discussion}}
\label{sec:discussion}
There are a few places where \mEdhoc{} can be improved,
which we found during this work and communicated to the authors.
%
We discuss them below.
%
